---
title: "Clean Permit Addresses"
author: "Alan Jackson"
date: '2022-04-28'
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(postmastr)
detach("package:GeocodeHou", unload=TRUE)
library(GeocodeHou)

savepath <-  "/home/ajackson/Dropbox/Rprojects/Curated_Data_Files/Houston_Permits/"
df <- readRDS(paste0(savepath, "Permit_data.rds"))

knitr::opts_chunk$set(echo = TRUE)
```

##  Start with manual data cleanup

Data had many issues (especially with what amounts to comments being placed
in the street address field).
So let's try to clean up the worst offenders.

```{r manual clean}

#   Clean up extra whitespace

df <- df %>% mutate(Address=str_squish(Address))

#   Let's run some tests

foo <- string_replace(df, "Address", "^(\\d+)\\s?\\d/\\d ", "\\1 ")
foo <- string_replace(df, "Address", " F M ", " FM ")
foo <- string_replace(df, "Address", "&", "AND")
foo <- string_replace(df, "Address", " BLD [A-Z1-9]$", " ")
#foo <- string_replace(df, "Address", "BUFFALO SPEEDWAY", "BUFFALO SPWY")
foo <- string_replace(df, "Address", " \\(PVT\\)",  "-----")
foo <- string_replace(df, "Address", "\\(.+\\)", "")
foo <- foo %>% mutate(Address=str_squish(Address))
foo <- string_replace(df, "Address", " A AND B", "")
foo <- string_replace(df, "Address", " FL[R]?\\s?\\d+$", " ")
foo <- string_replace(df, "Address", " STE [A-Z0-9-\\.]+$", " ")
foo <- string_replace(df, "Address", " [\\w]+-[\\w]+$", "")
foo <- string_replace(df, "Address", "^(\\d+) ST ", "\\1 SAINT ")
foo <- string_replace(df, "Address", "^(\\d+ [NSEW]) ST ", "\\1 SAINT ")

foo <- string_replace(foo, "Address", "( DR) \\S+$", "\\1")



# Protect Nasa Rd 1, it is unique in having a number following RD

foo <- df %>% mutate(Address=str_replace(Address, "NASA RD 1", "NASA ROAD 1"))

foo <- foo %>% # head(5000) %>% 
#   get rid of extra white space
       mutate(Address=str_squish(Address)) %>% 
#   get rid of address like 325 1/2. Get rid of the fraction
       string_replace( "Address", "^(\\d+)\\s?\\d/\\d ", "\\1 ", "Apply") %>% 
#   Collapse farm roads 
       string_replace("Output", " F M ", " FM ", "Apply") %>% 
#   change & to AND
       string_replace("Output", "&", "AND", "Apply") %>% 
#   Remove BLD (Building) designator
       string_replace("Output", " BLD\\s?[\\w]+$", "", "Apply") %>% 
#   Take care of Buffalo Speedway
       string_replace("Output", "BUFFALO SPEEDWAY", "BUFFALO SPWY", "Apply") %>% 
#   streets designated are Private are special and hard to geocode.
#   We will remove that designation, but then flag it in a comments column
  mutate(Special=ifelse(str_detect(Address, "\\(PVT\\)"), "Private", "")) %>% 
#   Remove anything in parenthesis
       string_replace("Output", "\\(.+\\)", "", "Apply") %>% 
#   Remove "A AND B"
       string_replace("Output", " A\\s?AND\\s?B", "", "Apply") %>% 
#   Remove "FL" for Floor
       string_replace("Output", " FL[R]?\\s?\\w+$", " ", "Apply") %>% 
#   Remove "STE" for Suite
       string_replace("Output", " STE [A-Z0-9-\\.]+$", " ", "Apply") %>% 
#   Remove "." but carefully. Collapse extra white space afterwards
       mutate(Output=str_replace_all(Output, "\\.", " ")) %>% 
       mutate(Output=str_squish(Output)) %>%  
#   Remove stuff with dashes in them except John-A
       mutate(Output=str_replace(Output, "JOHN-A", "JOHN:A")) %>% 
       string_replace("Output", " [\\w]+-[\\w]+$", "", "Apply") %>% 
       mutate(Output=str_replace(Output, "JOHN:A", "JOHN-A")) %>% 
#   Convert ST to SAINT
       string_replace("Output", "^(\\d+) ST ", "\\1 SAINT ", "Apply") %>% 
       string_replace("Output", "^(\\d+ [NSEW]) ST ", "\\1 SAINT ", "Apply") %>% 
#   Correct Parkway abbreviation
       string_replace("Output", " PKY", " PKWY", "Apply") %>% 
#   Correct Plaza streets
       string_replace("Output", 
                      "(LAWNDALE|LOVE|GREENWAY|MEYERLAND|WILCREST) PLAZA", 
                      "\\1 PLZ", "Apply") %>% 
#   Clean up stuff added to AVENUE
     string_replace("Output", "AVENUE ([A-Z]{1}) .+", "AVENUE \\1", "Apply") %>%   
#   Clean up stuff added to Riverway
       string_replace("Output", "RIVERWAY \\d+", "RIVERWAY", "Apply") %>%  
#   Clean up Sam Houston Pkwy
       string_replace("Output", "SAM HOUSTON NORTH", "SAM HOUSTON", "Apply") %>%  
       string_replace("Output", "SAM HOUSTON SOUTH", "SAM HOUSTON", "Apply") %>%  
       string_replace("Output", "SAM HOUSTON EAST",  "SAM HOUSTON", "Apply") %>%  
       string_replace("Output", "SAM HOUSTON WEST",  "SAM HOUSTON", "Apply") %>%   
#   Repair staff sgt macario
       string_replace("Output", "S/SGT MACARIO", "SSGT MACARIO", "Apply") %>%  
#   Repair avenida
       string_replace("Output", "AMERICAS .+$", "AMERICAS", "Apply") %>% 
#   Repair el camino real
       string_replace("Output", "EL CAMINO REAL .+$", "EL CAMINO REAL", "Apply") %>% 
#   Repair LOOP
       string_replace("Output", "NORTH LOOP .+$", "NORTH LOOP FWY", "Apply") %>% 
       string_replace("Output", "SOUTH LOOP .+$", "SOUTH LOOP FWY", "Apply") %>% 
       string_replace("Output", "EAST LOOP .+$", "EAST LOOP FWY", "Apply") %>% 
       string_replace("Output", "WEST LOOP .+$", "WEST LOOP FWY", "Apply") 

Types <- paste(c(" ALY", "AVE", "BLF", "BLVD", "BND", "BR", "BRG", "BRK", 
                 "BYP", "CIR",  "COR", "CRES", "CRK", "CRST", "CT", "CTR",
                 "CTS", "CV", "CYN", "DL", "DR",  "DRS", "EST", "ESTS",
                  "EXT", "FLD", "FLS", "FRK", "FRST", "FRY", "FWY", "GDN", 
                  "GDNS", "GLN", "GRN", "GRV", "HBR", "HL", "HLS", "HOLW",
                  "HTS", "HVN", "HWY",  "INLT", "IS", "KNL", "KNLS", "LDG",
                  "LK", "LKS", "LN", "LNDG", "MDW", "MDWS",
                  "ML", "MNR", "MT", "MTN", "PATH", "PKWY",
                  "PL",  "PLZ", "PNE", "PNES", "PR", "PRT", "PSGE", "PT",
                  "RD", "RDG", "RIV", "RST", "SHR", "SHRS",
                  "SPG", "SPUR", "SPWY", "SQ", "ST", "STA", "STRM", "TER",
                  "TRCE", "TRL", "VIS", "VL", "VLG", "VLY", "VW", "WALK",
                  "WLS",  "XING"), collapse="| ")
# 
 foo <- foo %>% 
   string_replace("Output", paste0("( ",Types,") .+$"), "\\1", "Apply")  

 More_types <- " MALL| WAY| PLACE"
 
 foo <- foo %>% 
   string_replace("Output", paste0("(",More_types,") \\w?\\d+$"), "\\1", "Apply")  
 
 
Farm_roads <- "FM 1960|FM 2100|FM 2351|FM 529|FM 2920|FM 1485|FM 2855|FM 1093|FM 2234|FM 362|FM 1942|FM 1314|FM 1463|FM 723|FM 1464|FM 686|FM 2978|FM 528|FM 521 |FM 1098|FM 149|FM 1959|FM 359|FM 1488|FM 249|FM 2917|FM 1736|FM 526"

foo <- foo %>% 
  string_replace("Output", paste0("(",Farm_roads,") .+$"), "\\1", "Apply")

# Unprotect Nasa Rd 1, it is unique in having a number following RD

foo <- foo %>% mutate(Output=str_replace(Output, "NASA ROAD 1", "NASA RD 1"))
#     What are ambiguous suffixes?

All_types <- df_names_only$Street_type %>% sort() %>% unique() 
  
All_types <- All_types[2:length(All_types)] # get rid of blank

All_types <- paste0(" ", All_types)

All_types <- paste0(All_types, collapse="|")

foobar <- df_names_only %>% 
  group_by(Street_name) %>% 
    summarize(last(Street_name)) %>% 
  mutate(Hit=str_detect(Street_name, All_types))

```


##      Parse address data

```{r parse}

pm_dictionary(type = "state", filter = c("TX"), case = "title", locale = "us")
dirs <- pm_dictionary(type = "directional", 
                      filter = c("N", "S", "E", "W"), locale = "us")
TX <- pm_dictionary(type = "state", filter = "TX", 
                    case = c("title", "upper"), locale = "us")

hou <- pm_append(type="suffix", input=c("SPEEDWAY", "SPWY", "PLACE"), 
                 output=c("SPWY", "SPWY", "PL"))
Type_dict <- pm_dictionary(type="suffix", append=hou, case="upper")

foo <- pm_identify(foo, var="Output") # add ID fields

foo2 <- pm_prep(foo, var="Output", type="street") # Prep data

foo2 <- pm_houseFrac_parse(foo2)

foo2 <- pm_house_parse(foo2)

foo2 <- pm_streetDir_parse(foo2, dirs)

foo2 <- pm_streetSuf_parse(foo2, Type_dict)
 
foo2 <- pm_street_parse(foo2)

foo2 <- foo2 %>% 
  mutate(pm.street=str_replace(pm.street, " 1 At 2", " 1/2")) 

foo2 <- foo2 %>% 
  mutate(pm.street=str_to_upper(pm.street)) %>% 
  mutate(pm.streetSuf=str_to_upper(pm.streetSuf)) %>% 
  mutate(pm.preDir=replace_na(pm.preDir, "")) %>% 
  mutate(pm.streetSuf=replace_na(pm.streetSuf, ""))

foo <- pm_replace(foo2, source=foo)

```

##        Now let's geocode

```{r geocode}

Exact_match <- NULL
Failed_match <- NULL

for (i in 1:nrow(foo)){ # first look for exact matches
#for (i in 2500:5000){ # first look for exact matches takes about 1 minute
  if (i%%1000==0) { print(paste("---",i,"---")) }
  tmp <- match_exactly(foo[i,]$pm.house, foo[i,]$pm.preDir, foo[i,]$pm.street,
                       foo[i,]$pm.streetSuf, foo[i,]$Zipcode)
  if (tmp$Success){ #   success
    Exact_match <- cbind(foo[i,], tmp) %>% 
      select(pm.id, pm.house, pm.preDir, pm.street, pm.streetSuf, Zipcode, Lat, Lon) %>% 
      rbind(., Exact_match)
  } else { #  Fail exact match
    Failed_match <- cbind(foo[i,], tmp) %>% 
      select(pm.id, pm.house, pm.preDir, pm.street, pm.streetSuf, Zipcode, Fail,
             Lat, Lon) %>% 
      rbind(., Failed_match)
  }
}

saveRDS(Exact_match, paste0(savepath, "Keep_Exactmatch.rds"))
saveRDS(Failed_match, paste0(savepath, "Keep_Failedmatch.rds"))

```

##  Now let's try to repair the failures


```{r bad zip}

Failed_match <- Failed_match %>% mutate(Correction=NA)

for (i in 1:nrow(Failed_match)){
  if (i%%100==0) { print(paste("---",i,"---")) }
  target <- Failed_match[i,]
  tmp <- repair_zipcode(target$pm.house, target$pm.preDir, target$pm.street,
                       target$pm.streetSuf, target$Zipcode)
  if (tmp$Success){ #   success
    Failed_match[i,]$Lat <- tmp$Lat
    Failed_match[i,]$Lon <- tmp$Lon
    Failed_match[i,]$Fail <- paste(Failed_match[i,]$Fail, "Zipcode")
    Failed_match[i,]$Correction <- tmp$New_zipcode
  } 
}

```

##    Look for bad names

```{r bad names}

for (i in 1:nrow(Failed_match)){
  if (Failed_match[i,]$Fail!="Street_name") {next} # skip if name isn't the issue
  if (Failed_match[i,]$Lat > 0) {next} # skip if zipcode resolved it
  if (i%%10==0) { print(paste("---",i,"---")) }
  target <- Failed_match[i,]
  tmp <- repair_name(target$pm.house, target$pm.preDir, target$pm.street,
                       target$pm.streetSuf, target$Zipcode)
  if (tmp$Success){ #   success
    Failed_match[i,]$Lat <- tmp$Lat
    Failed_match[i,]$Lon <- tmp$Lon
    Failed_match[i,]$Correction <- tmp$New_name
  } else {
    Failed_match[i,]$Fail <- paste("Street_name",tmp$Fail)
  }
}

```

##      Bad types

```{r bad types}

for (i in 1:nrow(Failed_match)){
  if (Failed_match[i,]$Fail!="Street_type") {next} # skip if type isn't the issue
  if (i%%10==0) { print(paste("---",i,"---")) }
  target <- Failed_match[i,]
  tmp <- repair_type(target$pm.house, target$pm.preDir, target$pm.street,
                       target$pm.streetSuf, target$Zipcode)
  if (tmp$Success){ #   success
    Failed_match[i,]$Lat <- tmp$Lat
    Failed_match[i,]$Lon <- tmp$Lon
    Failed_match[i,]$Correction <- tmp$New_type
  } else {
    Failed_match[i,]$Fail <- paste("Street_type",tmp$Fail)
  }
}

```

##     Bad numbers


```{r bad numbers}


for (i in 1:nrow(Failed_match)){
  if (Failed_match[i,]$Fail!="Street_num") {next} # skip if name isn't the issue
  if (i%%10==0) { print(paste("---",i,"---")) }
  target <- Failed_match[i,]
  tmp <- repair_number(target$pm.house, target$pm.preDir, target$pm.street,
                       target$pm.streetSuf, target$Zipcode)
  if (tmp$Success){ #   success
    Failed_match[i,]$Lat <- tmp$Lat
    Failed_match[i,]$Lon <- tmp$Lon
    Failed_match[i,]$Correction <- tmp$Result
  } else {
    Failed_match[i,]$Fail <- paste("Street_num",tmp$Fail)
  }
}

```


##        Bad prefix


```{r bad prefix}

for (i in 1:nrow(Failed_match)){
  if (Failed_match[i,]$Fail!="Prefix") {next} # skip if name isn't the issue
  if (i%%10==0) { print(paste("---",i,"---")) }
  target <- Failed_match[i,]
  tmp <- repair_prefix(target$pm.house, target$pm.preDir, target$pm.street,
                       target$pm.streetSuf, target$Zipcode)
  if (tmp$Success){ #   success
    Failed_match[i,]$Lat <- tmp$Lat
    Failed_match[i,]$Lon <- tmp$Lon
    Failed_match[i,]$Correction <- tmp$New_prefix
  } else {
    Failed_match[i,]$Fail <- paste("Prefix",tmp$Fail)
  }
}


```

##        Results

Very nearly half of the failures were recovered. That is pretty good. So
out of 130,000 addresses (with many duplicates), all but 4400 were geocoded.

```{r near final}

#   Combine all the files and produce a final failure file that we can try a few
#   more tricks with.

saveRDS(Exact_match, paste0(savepath, "Keep_Exactmatch_final.rds"))
saveRDS(Failed_match, paste0(savepath, "Keep_Failedmatch_final.rds"))

Final <- left_join(foo, select(Exact_match, pm.id, Lat, Lon), by="pm.id")

Final <- left_join(Final,
                   select(Failed_match, pm.id, Lat, Lon, Fail, Correction),
                   by="pm.id")

Final <- Final %>% 
  rowwise() %>% 
  mutate(Lat.x=ifelse(is.null(Lat.x), Lat.y, Lat.x)) %>% 
  mutate(Lon.x=ifelse(is.null(Lon.x), Lon.y, Lon.x)) %>% 
  rename(Lon=Lon.x, Lat=Lat.x) %>% 
  select(-Lat.y, -Lon.y)

#   Prefix

Final <- Final %>% 
  mutate(pm.preDir=case_when(
    replace_na(str_detect(Fail, "^Prefix$"), FALSE) ~ Correction,
    TRUE                                            ~ pm.preDir
  ))

#   Name

Final <- Final %>% 
  mutate(pm.street=case_when(
    replace_na(str_detect(Fail, "^Street_name$"), FALSE) ~ Correction,
    TRUE                                                 ~ pm.street
  ))

#   Zipcode

Final <- Final %>% 
  mutate(Zipcode=case_when(
    replace_na(str_detect(Fail, "^Street_name Zipcode$"), FALSE) ~ Correction,
    TRUE                                                         ~ Zipcode
  ))

#   Type

Final <- Final %>% 
  mutate(pm.streetSuf=case_when(
    replace_na(str_detect(Fail, "^Street_type$"), FALSE) ~ Correction,
    TRUE                                                 ~ pm.streetSuf
  ))

#   Rename some columns to better document

Final <- Final %>% 
  rename(Number=pm.house, Prefix=pm.preDir, Name=pm.street, Type=pm.streetSuf,
         Original_address=Address, Cleaned_address=Output)

saveRDS(Final, paste0(savepath, "Clean_Final_2017_to_Mar_2022.rds"))


```

```{r}

#     Let's do a little repair until whole thing is rerun.


Failed_match <- readRDS(paste0(savepath, "Keep_Failedmatch_final.rds"))

Failed_match[str_detect(replace_na(Failed_match$Correction, ""),
                        "^\\d{5}$"),]$Fail <- 
  "Street_name Zipcode"
           


```










