---
title: "Gun Permit Data"
author: "Alan Jackson"
date: "7/3/2021"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(tabulizer)
library(textreadr)

output_folder <- "/home/ajackson/Dropbox/Rprojects/Curated_Data_Files/Gun_Permits/"

#   read in standard county name file

Counties <- readRDS("/home/ajackson/Dropbox/Rprojects/Curated_Data_Files/Texas_Counties/County_Names.rds")

options(stringsAsFactors = FALSE)

knitr::opts_chunk$set(echo = TRUE)
```

## Here be files by zipcode, with varying names...

https://www.dps.texas.gov/sites/default/files/documents/rsd/ltc/reports/2016calendar/byzipcode/17licenseapplicationsissued.pdf

https://www.dps.texas.gov/sites/default/files/documents/rsd/ltc/reports/2020calendar/byzipcode/17licenseapplicationsissued.pdf

Earlier files already downloaded a few years ago.

```{r read by zip, eval=FALSE}
#   Download all the pdfs

###output_folder <- "~/Dropbox/Rprojects/WilsonCountyGunStudy/GunPermits"
for (year in 2017:2020) {
  url <- paste0("http://www.dps.texas.gov/sites/default/files/documents/rsd/ltc/reports/",year,"calendar/byzipcode/17licenseapplicationsissued.pdf")

  outfile <- paste0("LicenseApplicationsIssued_by_Zip",year,".pdf")
  m <- download(url, output_folder, outfile) 
  print(m)
}

```

## Here be files by county, with varying names...

https://www.dps.texas.gov/sites/default/files/documents/rsd/ltc/reports/2020fiscal/bycounty/25licenseapplicationsissued.pdf

```{r read by county, eval=FALSE}
#   Download all the pdfs

for (year in 2017:2020) {
  url <- paste0("http://www.dps.texas.gov/sites/default/files/documents/rsd/ltc/reports/",year,"calendar/bycounty/25licenseapplicationsissued.pdf")
  outfile <- paste0("LicenseApplicationsIssued_by_County",year,".pdf")
  m <- download(url, output_folder, outfile) 
  print(m)
}

```

##  Parse the County PDF's

These are for 2017-2020, which I did not already have done. The third page
of the 2020 pdf refuses to parse nicely, so I extracted the values using
Acrobat, and manually created an approrpiate CSV file from them. So it goes.


```{r parse county pdfs}

parse_county <- function(path, year){
  options(stringsAsFactors = FALSE)
  
  final <- data.frame(Year=character(),
                      County=character(),
                      Applicants=numeric(),
                      Percent=numeric(),
                      stringsAsFactors = FALSE)
  
  a <- extract_tables(path)
  skip <- 3
  tnum <- 1:4  # table numbers to grab
  if (year=="2020") {tnum <- 1:2} # third page a problem
  
  for (i in tnum){
    temp <- as.data.frame(a[[i]][skip:nrow(a[[i]]),1:3], stringsAsFactors = FALSE)
    names(temp) <- c("County", "Applicants", "Percent")
    temp <- cbind(Year=as.character(year), temp)
    temp$Applicants <- as.numeric(gsub(",","",temp$Applicants))
    temp$Percent <- as.numeric(gsub("%","",temp$Percent))
    final <- bind_rows(final, temp)
    
    if (i <4){
      temp <- as.data.frame(a[[i]][skip:nrow(a[[i]]),4:6], stringsAsFactors = FALSE)
      names(temp) <- c("County", "Applicants", "Percent")
      temp <- cbind(Year=as.character(year), temp)
      temp$Applicants <- as.numeric(gsub(",","",temp$Applicants))
      temp$Percent <- as.numeric(gsub("%","",temp$Percent))
      final <- bind_rows(final, temp)
    }
  }
  return(final)
}

final <- data.frame(Year=character(),
                    County=character(),
                    Applicants=numeric(),
                    Percent=numeric(),
                    stringsAsFactors = FALSE)

path <- paste0(output_folder,"LicenseApplicationsIssued_by_County")
for (year in c(2017:2020)) {
  print(paste0("parse ",year))
  a <- parse_county(paste0(path,year,".pdf"), year)
  final <- bind_rows(final, a)
}

#   Read in manually constructed csv file and append

foo <- read_csv(paste0(output_folder, "County2020_page_3.txt"), 
                col_names=FALSE) %>% 
  rename(County=X1, Applicants=X2, Percent=X3) %>% 
  add_column(Year="2020", .before=1)

final <- bind_rows(final, foo)

#     enforce some consistency

final$County <- str_replace(final$County, "\\[Out of State County\\]", "NONE")
final$County <- str_replace(final$County, "\\[NONE\\]", "NONE")

#   Clean up the county names 

final <- final %>% 
  mutate(County=str_to_title(County))

final[final$County=="Dewitt",]$County <- "DeWitt"  # yeah, why do they spell it that way?

#     Save to a file

saveRDS(final, paste0(output_folder, "Licenses_by_County_2017to2020.rds"))


```


##   Do a little QC on the data

```{r QC step}

summary(final)

final %>%
  filter(Percent<25) %>%
  filter(County!="Total") %>%
  ggplot(aes(x=Applicants, y=Percent)) +
  geom_point()

#   make a rotated bar chart of applications by county
final %>%
  filter(County!="Total") %>%
  ggplot(aes(x=County, y=Applicants)) +
  geom_col() +
  coord_flip()

```

##    Parse zipcode gun permit files

```{r parse zipcode files}

parse_zip <- function(path, year){
  options(stringsAsFactors = FALSE)
  
  final <- data.frame(Year=character(),
                      Zipcode=character(),
                      Applicants=numeric(),
                      stringsAsFactors = FALSE)
  
  a <- extract_tables(path)
  skip <- 3
  tnum <- 1:length(a)  # table numbers to grab
  
  for (i in tnum){
    print(paste("----", i, "-----"))
    if (dim(a[[i]])[1] < 40 && year == 2012) {next}
    if (i%%2==1 && year == 2018) {next}
    width <- dim(a[[i]])[2]
    temp <- as.data.frame(a[[i]][skip:nrow(a[[i]]),1:width], stringsAsFactors = FALSE)
    # split table vertically
    if (width==8){
      temp <- temp %>% bind_rows(rename(select(.,V3,V4), V1=V3, V2=V4), rename(select(.,V5,V6),V1=V5, V2=V6),rename(select(.,V7,V8), V1=V7, V2=V8))%>%
      select(V1, V2)
  }
  else if (width==6){
      temp <- temp %>% bind_rows(rename(select(.,V3,V4), V1=V3, V2=V4), rename(select(.,V5,V6),V1=V5, V2=V6))%>%
      select(V1, V2)
  }
  else if (width == 4) {
      temp <- temp %>% bind_rows(rename(select(.,V3,V4), V1=V3, V2=V4))%>%
      select(V1, V2)
  }
    
    names(temp) <- c("Zipcode", "Applicants")
    temp <- cbind(Year=as.character(year), temp)
    temp$Applicants <- as.numeric(gsub(",","",temp$Applicants))
    final <- bind_rows(final, temp) %>%
      na.omit()
    
    
  }
  
  return(final)
} ####   end of function

finalzip <- data.frame(Year=character(),
                    Zipcode=character(),
                    Applicants=numeric(),
                    stringsAsFactors = FALSE)

path <- paste0(output_folder,"LicenseApplicationsIssued_by_Zip")
for (year in c(2017:2020)) {
  print(paste0("parse ",year))
  a <- parse_zip(paste0(path,year,".pdf"), year)
  finalzip <- bind_rows(finalzip, a)
}

#   drop rows with zipcodes not in Texas

finalzip$Zipcode <- as.numeric(finalzip$Zipcode)
finalzip <- finalzip %>% 
  na.omit() %>%
  filter(Zipcode==73301 | 
         between(Zipcode,75001,75501) |
         between(Zipcode,75503,79999))
  

#     Save to a file

saveRDS(final, paste0(output_folder, "Licenses_by_Zip_2017to2020.rds"))

```

	
##   Do some somple QC on the zipcode data

```{r qc zipcode data}

finalzip %>%
  filter(Year != "1997")%>% # low number for 1st year
  filter(Year != "1999")%>% # low numbers for unknown reason
  group_by(Zipcode) %>%
  summarize(dev=sd(Applicants), avg=mean(Applicants), minimum=min(Applicants), maximum=max(Applicants)) %>%
  arrange(desc(avg))


```

