---
title: "Houston Neighborhoods"
author: "Alan Jackson"
date: '2022-05-07'
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(sf)

path <- "/home/ajackson/Dropbox/Rprojects/Curated_Data_Files/Neighborhoods/"

googlecrs <- 4326

knitr::opts_chunk$set(echo = TRUE)
```

##    Trials and tribulations

*** Add Neighborhoods. ***
Neighborhoods are a difficult dataset. The definition is rather loose, and they may overlap. And the sources for neighborhood data are limited. I looked at three: Zillow, City of Houston GIS, and Wikimapia. Zillow is very nice data, easy to download and work with, but not complete. I think they disallow overlapping polygons. The City of Houston is really subdivisions, and is a work in progress, so it didn't help much. So Wikimapia. The data itself is excellent. For some points, as many as 5 associated neighborhoods, so it seems to be pretty complete. Plus, if I find errors, I can edit the source data.
However, many challenges downloading and then loading the data into R.

To download, I used the Wikimapia API builder, noted in http://blog.cartong.org/2014/10/30/tutorial-using-wikimapia-data-in-qgis/
which generated a command line of

http://wikimapia.org/api?action=create_key

for ajackson.org
23523C7E-AEAF8DB3-CDE1EFEC-2B447357-9E16F369-A1924E0E-878711F2-1F552310

for adelieresources.com
23523C7E-F35339CC-105CD5F5-57788CE9-1CE8B9EE-AE7815B5-4A8DC5BE-19CC32CE

http://api.wikimapia.org/?key=MYKEY-GETYOUROWN&function=box&coordsby=bbox&bbox=-95.789963%2C29.518566%2C-95.005814%2C30.005814&category=4621&count=2000&format=kml

https://www.im-portal.org/

I generated JSON, KML, and XML outputs. I struggled for hours to try to read
*any* of the three into R, without success. Finally I got a clue and found
success by going outside R and using ogr. The Wikimapia files appear to be
malformed somehow, but ogr can handle it.

ogr2ogr -f GML Neighborhoods.gml Neighborhoods.kml -explodecollections

ogr2ogr -f "ESRI Shapefile" Nbhd_point.shp Neighborhoods.gml -sql "select * from Layer0 where OGR_GEOMETRY='POINT'" -lco SHPT=POINT

ogr2ogr -f "ESRI Shapefile" Nbhd_line.shp Neighborhoods.gml -sql "select * from Layer0 where OGR_GEOMETRY='LINESTRING'" -lco SHPT=ARC


```{r read in}

#   Read in the shapefiles we created with ogr2ogr, to a simple feature
Neighborhoodpolys <- sf::read_sf(paste0(path, "/Wikimapia_2022/Nbhd_line.shp"))

# now need to delete extra crap from file, just to be tidy
temp <- Neighborhoodpolys %>% select(Name, descriptio, geometry)
summary(temp) # check to see what it looks like. Could also do plot(temp)
#     Name            descriptio                      geometry  
# Length:975         Length:975         GEOMETRYCOLLECTION:975  
# Class :character   Class :character   epsg:4326         :  0  
# Mode  :character   Mode  :character   +proj=long...     :  0  

#   set CRS. Since Wikimapia is based on Google maps photos, CRS is 
#   same as Google.
sf::st_crs(temp) <- googlecrs
#   turn lines into polygons
###temp <- sf::st_polygonize(temp)
###temp <- sf::st_as_sf(temp)

# But first fix a few issues
temp$Name <- str_replace(temp$Name,"&#039;","'")
temp$Name <- str_replace(temp$Name,"Associaition","Association")
temp$Name <- str_replace(temp$Name,"&amp;","&")

#     Add short names for better labels

temp2 <- temp %>% 
  mutate(Short_name=str_remove(Name, "\\(?Super Neighborhood \\d+\\)?")) %>%  
  mutate(Short_name=str_remove(Short_name, "\\(?SN ?\\d+\\)?")) %>% 
  mutate(Short_name=str_remove(Short_name, "\\(.*\\)")) %>% 
  mutate(Short_name=str_remove(Short_name, " Subdivision")) %>% 
  mutate(Short_name=str_remove(Short_name, " [Nn]eighborhood")) %>% 
  mutate(Short_name=str_replace(Short_name, " Section", " Sec")) %>% 
  mutate(Short_name=str_remove(Short_name, "\\(South Briar Community Association")) %>% 
  select(Name, Short_name, descriptio, geometry)
#   turn lines into polygons
###temp2 <- sf::st_polygonize(temp2)

#   Save file

saveRDS(temp2, paste0(path, "Neighborhood_polys_wikimapia.rds"))

```

##        Let's make a nice map


```{r map them}

library(leaflet)

ggplot(temp2) + 
  geom_sf()  
  geom_sf_label(aes(label=Short_name), label.size=NA, alpha=0.0, size=2)

leaflet(temp2) %>% 
  leaflet::setView(lng = -95.3103, lat = 29.7752, zoom = 8 ) %>%   
  addTiles() %>% 
  addPolygons()

```


##        City of Houston super neighborhoods

Looks good, looks watertight and comprehensive. Possibly the most useful
set of neighborhoods for many purposes.

https://cohgis-mycity.opendata.arcgis.com/datasets/deae15e7f3db4b198946c17f0d169c39/explore?location=29.721674%2C-95.357101%2C10.99

Download the shapefile


```{r COH hoods}

#   Read in the shapefiles 
Neighborhoodpolys <- sf::read_sf(paste0(path,
                    "/CityOfHouston_2022/COH_SUPER_NEIGHBORHOODS.shp"))

temp <- Neighborhoodpolys %>% select(SNBNAME, geometry)
summary(temp) # check to see what it looks like. Could also do plot(temp)
#  SNBNAME                   geometry 
# Length:88          POLYGON      :88  
# Class :character   epsg:4326    : 0  
# Mode  :character   +proj=long...: 0  
st_crs(temp) <- googlecrs

temp <-temp %>% rename(Name=SNBNAME)

saveRDS(temp, paste0(path, "Neighborhood_polys_COH.rds"))


leaflet(temp) %>% 
  leaflet::setView(lng = -95.3103, lat = 29.7752, zoom = 8 ) %>%   
  addTiles() %>% 
  addPolygons()

```



##        Harris County city boundaries

Looks good, looks watertight and comprehensive. Possibly the most useful
set of neighborhoods for many purposes.

From Kinder Institute,
Harris County Appraisal District. (2020). Boundaries of Cities in Harris County.
[Data set]. Rice University-Kinder Institute: UDP. https://www.kinderudp.org/#/datasetCatalog/3xmql87o2y1v 


```{r County cities}

#   Read in the shapefiles 
CityPolys <- sf::read_sf(paste0(path,
                    "/CitiesInCounty/City.shp"))

temp <- CityPolys %>% select(name, geometry)
summary(temp) # check to see what it looks like. 
 #     name                    geometry 
 # Length:34          MULTIPOLYGON :34  
 # Class :character   epsg:NA      : 0  
 # Mode  :character   +proj=lcc ...: 0  

#   Data is projected NAD_1983, Lambert

#   So let's convert to lat long

fgdb <- paste0(path, "CitiesInCounty/City.dbf")

# Read the feature class
temp <- sf::st_read(fgdb)

#   unproject the zipcode data to lat longs on proper datum

templatlong <- st_transform(temp, googlecrs)
summary(templatlong)
 #     code               name             GlobalID           Shape_area       
 # Length:34          Length:34          Length:34          Min.   :1.816e+06  
 # Class :character   Class :character   Class :character   1st Qu.:3.761e+07  
 # Mode  :character   Mode  :character   Mode  :character   Median :5.833e+07  
 #                                                          Mean   :6.691e+08  
 #                                                          3rd Qu.:1.610e+08  
 #                                                          Max.   :1.704e+10  
 #   Shape_len                geometry 
 # Min.   :   9867   MULTIPOLYGON :34  
 # 1st Qu.:  32456   epsg:4326    : 0  
 # Median :  41616   +proj=long...: 0  
 # Mean   : 191784                     
 # 3rd Qu.:  79534                     
 # Max.   :3567219                     

saveRDS(templatlong, paste0(path, "City_polys.rds"))


leaflet(templatlong) %>% 
  leaflet::setView(lng = -95.3103, lat = 29.7752, zoom = 8 ) %>%   
  addTiles() %>% 
  addPolygons()

```













