---
title: "Texas Death Data 2006-2017"
author: "Alan Jackson"
date: "6/15/2021"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)

path <- "/home/ajackson/Dropbox/Rprojects/Curated_Data_Files/Texas_Deaths/"

pop_path <- "/home/ajackson/Dropbox/Rprojects/Curated_Data_Files/Texas_Demographics/"
Pop <- readRDS(paste0(pop_path, "County_Age_Sex.rds")) %>% 
  select(FIPS, CNTY_NM, Total_pop) %>% 
  rename(County=CNTY_NM)

knitr::opts_chunk$set(echo = TRUE)
```

## Texas Health Department Death Data

Load, check, and clean up [data downloaded](https://healthdata.dshs.texas.gov/dashboard/births-and-deaths/deaths-2006-2017)
from the state of Texas health
department website on deaths for the period 2006-2017 (which is what is
available as of this date)

To get all the data, it will need to be downloaded many times. Once for the 
Age distribution, once for sex, once for ethnicity, and once for the total
per county in hopes of recovering some of the hidden values. All of these
will eventually get combined. And they must each be downloaded year-by-year.
Also download the yearly state totals, and the total state totals for 
imputing the censored values.

Each file is downloaded as what they call a "crosstab", which is just a
CSV file.

Values less than some magic minimum are converted to "---" in the file,
while zero is zero. It appears that that magic minimum is 9. But we'll 
test that.

To begin, we will read in and examine the full state totals.

```{r state totals}



# Squirrelly encoding seems to flummox read_csv
#     "---" is the left-censored counts flag

###########     Read in totals for all years and the state

Df_State_tot <- read.delim(paste0(path, "Input_data/State_allyears_deaths.csv"),
                   sep="\t",
                   fileEncoding="UTF-16LE",
                   na.strings="---",
                   skip=1,
                   header=FALSE,
                   colClasses="character",
                   stringsAsFactors = FALSE
                   ) %>% 
  rename("County"=V1, "Condition"=V2, "Deaths"=V3) %>%  
  #   F&^%ers put commas in the numbers
  mutate_at(vars(contains("Deaths")), 
         funs(str_remove_all(., ","))) %>% 
  #   Make numbers numbers
  type_convert(col_types = paste0("cc", strrep("i",1)))

###########     Read in totals for each year and the state

Df_State_yr_tot <- 
  data.frame(year=2006:2017) %>% 
  mutate(filename=paste0(path, "Input_data/State_", year, "_deaths.csv")) %>% 
  mutate(file_contents=map(filename,
    ~ read.delim(.,
                   sep="\t",
                   fileEncoding="UTF-16LE",
                   na.strings="---",
                   skip=1,
                   header=FALSE,
                   colClasses="character",
                   stringsAsFactors = FALSE
                   ))) %>% 
  select(-filename) %>% 
    unnest() %>% 
  rename("County"=V1, "Condition"=V2, "Deaths"=V3) %>%  
  #   F&^%ers put commas in the numbers
  mutate_at(vars(contains("Deaths")), 
         funs(str_remove_all(., ","))) %>% 
  #   Make numbers numbers
  type_convert(col_types = paste0("icc", strrep("i",1)))

#   Impute values where possible

DF_Impute_Deaths_State_yr <- 
left_join(Df_State_yr_tot, Df_State_tot, by="Condition") %>% 
  rename(Deaths=Deaths.x, Tot_deaths=Deaths.y) %>% 
  select(-County.x, -County.y) %>% 
  group_by(Condition) %>% 
    mutate(Sum_na=sum(is.na(Deaths)),
           Sum=sum(Deaths, na.rm=TRUE),
           Impute_deaths=Deaths) %>% 
  ungroup() %>% 
  mutate(Impute_deaths=ifelse(is.na(Impute_deaths),
                              (Tot_deaths-Sum)/Sum_na,
                               Deaths)) 

```

Now county totals for all years


```{r county all years}

# Squirrelly encoding seems to flummox read_csv
#     "---" is the left-censored counts flag

###########     Read in By County

Df_county_all <- read.delim(paste0(path, "Input_data/County_allyears_deaths.csv"),
                   sep="\t",
                   fileEncoding="UTF-16LE",
                   na.strings="---",
                   skip=1,
                   header=FALSE,
                   colClasses="character",
                   stringsAsFactors = FALSE
                   ) %>% 
  rename("County"=V1, "Condition"=V2, "Deaths"=V3
         ) %>%  
  #   F&^%ers put commas in the numbers
  mutate_at(vars(contains("Deaths")), 
         funs(str_remove_all(., ","))) %>% 
  #   Make numbers numbers
  type_convert(col_types = paste0("cc", strrep("i",1))) %>% 
  #   Add county population number for later use in imputing
  left_join(., Pop, by="County")

#   Impute values where possible

# For each condition, look only at NA counties, sum up the population 
# of those counties, and then take the remaining imputed deaths from the
# state totals, divide by NA population, and then impute to each county
# that number times the county population.

smart.round <- function(x) {
  #   from stack overflow 
  #   https://stackoverflow.com/questions/32544646/round-vector-of-numerics-to-integer-while-preserving-their-sum/32544987#32544987
  #   re-integer vector components
  y <- floor(x)
  xx <- x[!is.na(x)]
  yy <- y[!is.na(y)]
  indices <- tail(order(xx-yy), round(sum(xx)) - sum(yy))
  yy[indices] <- yy[indices] + 1
  y[!is.na(y)] <-  yy
  
  y
}

set_min_one <- function(x){
  #   Take a vector of integer components that range from 0 up, and 
  #   steal from the larger values to set all the zeros to one. Used
  #   for imputing number of deaths.
  xx <- x[!is.na(x)]
  while((z=sum(xx==0))>0){
    twoplus <- min(sum(xx>1), z)
    if (twoplus==0){# something bad happened
      print("Too many zeros - something bad happened")
      print(xx)
      break
    }
    i_sub <- tail(order(xx), twoplus)
    i_add <- head(order(xx), twoplus)
    xx[i_sub] <- xx[i_sub] - 1
    xx[i_add] <- xx[i_add] + 1
  }
  #   Now steal from the larger values to push them down to 9, and
  #   redistribute those numbers to the smaller values.
  
  redist <- sum(pmax(0,xx-9)) # amount to redistribute
  # Set values >9 to 9
  xx <- pmin(9, xx)
  i <- 1
  while(redist>0) {
    #  Add 1 to a value, biased towards values that are already larger
    if ((runif(1,min=1, max=9) < xx[i]) &
        xx[i]<9) {
      xx[i] <- xx[i] + 1
      redist <- redist - 1
    }
    i <- ((i) %% length(xx)) + 1
  }
  
  #   Re-merge original and imputed values
  x[!is.na(x)] <-  xx
  
  x
}

DF_Impute_Deaths_County <- 
left_join(Df_county_all, Df_State_tot, by="Condition") %>% 
  rename(Deaths=Deaths.x, Tot_deaths=Deaths.y) %>% 
  select(-County.y) %>% 
  group_by(Condition) %>% 
    mutate(Sum_pop=sum(is.na(Deaths)*Total_pop, na.rm=TRUE),
           Sum_known=sum(Deaths, na.rm=TRUE),
           Max_pop=max(Total_pop[is.na(Deaths)]),
           ) %>% 
  ungroup() %>% 
  # First impute based on relative county population
  mutate(Impute_deaths=ifelse(is.na(Deaths),
                              (Tot_deaths-Sum_known)*Total_pop/Sum_pop,
                               Deaths)) %>% 
  #   Assume any remaining NA's are such rare things that replacing with 1 
  #   is good. These are values where the state total is < 10.
  mutate(Impute_deaths=ifelse(is.na(Impute_deaths),
                              1,
                              Impute_deaths)) %>% 
  #   Now turn the weights into integers while preserving the sum
  group_by(Condition) %>% 
    mutate(Impute_deaths=ifelse(is.na(Deaths),
                                Impute_deaths,
                                NA)) %>% 
    mutate(Impute_deaths=smart.round(Impute_deaths)) %>% 
    #   Now replace zeros with ones by stealing from larger numbers
    #   Until the largest imputed number is 9
    mutate(Impute_deaths=set_min_one(Impute_deaths)) %>% 
    #   Add Deaths that are not imputed back into imputed deaths column
    mutate(Impute_deaths=ifelse(is.na(Deaths), Impute_deaths, Deaths)) %>% 
  ungroup() %>% 
  select(County.x, Condition, Deaths, FIPS, Total_pop, Impute_deaths) %>% 
  rename(County=County.x, County_pop=Total_pop)
  
saveRDS(DF_Impute_Deaths_County, 
        paste0(path, "Imputed_Deaths_by_County_Total.rds"))



```

Let's take a look at the data just to make sure it is okay

```{r data check 1}

DF_Impute_Deaths_County %>% 
  filter(Condition=="Emphysema") %>% 
  ggplot(aes(y=County, x=Impute_deaths)) + 
  geom_col()
  
DF_Impute_Deaths_County %>% 
  filter(Condition=="Emphysema") %>% 
  mutate(Death_per_cap=1.e5*Impute_deaths/County_pop) %>% 
  filter(Death_per_cap>50,
         Impute_deaths>5) %>% 
  mutate(sigma=1.e5*((Impute_deaths+3)/County_pop - Impute_deaths/County_pop)) %>% 
  ggplot(aes(y=reorder(County, Death_per_cap), x=Death_per_cap)) + 
  geom_col() + 
  geom_errorbar(aes(xmin=Death_per_cap-sigma, xmax=Death_per_cap+sigma))

DF_Impute_Deaths_County %>% 
  filter(is.na(Deaths)) %>% # only imputed values
  ggplot(aes(x=Impute_deaths)) +
  geom_bar()

DF_Impute_Deaths_County %>% 
  filter(is.na(Deaths)) %>% # only imputed values
  mutate(Impute_per_cap=1.e5*Impute_deaths/County_pop) %>% 
  ggplot(aes(x=Impute_per_cap)) +
  geom_histogram()

DF_Impute_Deaths_County %>% 
  filter(!is.na(Deaths)) %>% # only known values
  mutate(Impute_per_cap=1.e5*Impute_deaths/County_pop) %>% 
  ggplot(aes(x=Impute_per_cap)) +
  geom_histogram()

```

Now read in county data by year and condition

```{r county by year}

Df_County_yr <- 
  data.frame(year=2006:2017) %>% 
  mutate(filename=paste0(path, "Input_data/Deaths_by_county_", year,".csv")) %>% 
  mutate(file_contents=map(filename,
    ~ read.delim(.,
                   sep="\t",
                   fileEncoding="UTF-16LE",
                   na.strings="---",
                   skip=1,
                   header=FALSE,
                   colClasses="character",
                   stringsAsFactors = FALSE
                   ))) %>% 
  select(-filename) %>% 
    unnest() %>% 
  rename("County"=V1, "Condition"=V2, "Deaths"=V3) %>%  
  #   F&^%ers put commas in the numbers
  mutate_at(vars(contains("Deaths")), 
         funs(str_remove_all(., ","))) %>% 
  #   Make numbers numbers
  type_convert(col_types = paste0("icc", strrep("i",1)))

Df_County_yr <- Df_County_yr %>% 
  #   add minimum that sum over years must be
  group_by(County, Condition) %>% 
    mutate(Min_sum_deaths=sum(is.na(Deaths))) %>% 
  ungroup()

####   Let's backtrack a bit and fix the DF_Impute_Deaths_County
####   by applying he minimum values we just calculated to it.

foo <- Df_County_yr %>% 
  group_by(County, Condition) %>% 
    summarize(Min_sum_deaths=sum(is.na(Deaths)))  
  
DF_Impute_Deaths_County <- left_join(DF_Impute_Deaths_County, 
                                     foo, 
                                     by=c("County", "Condition"))
Fix_minimum <- function(df){
    x <- df[is.na(df$Deaths),] # pick out only rows to be updated
      # add row number so order can be recovered
    x <- x %>% mutate(ID=row_number()) %>% 
      # order by adjustment so we'll fix it in one pass
      arrange(adjust)
  print(paste("------", df[1,]$Condition))  
  #   If all values are NA, treat as special
  if (sum(is.na(x$Deaths))==nrow(x)) {
    print("I'm special")
    x$Impute_deaths <- x$Min_sum_deaths
    x <- x %>% arrange(ID) %>% 
      select(-ID)
    df[is.na(df$Deaths),] <- x # add back only rows updated
    return(df$Impute_deaths) # return for this case
  }
  redist <- -sum(pmin(0,x$adjust)) # amount to redistribute (neg values only)
  i <- 1
  while(redist>0) {
    #print(paste("-2-", i, redist))
    #  Add adjust to a value, or if >=0, steal from redist
    if (x$adjust[i]<0) {
    #print("-3-")
      x$Impute_deaths[i] <- x$Impute_deaths[i] - x$adjust[i]
      x$adjust[i] <- 0
    } else {
   #print("-4-")
      if (x$Impute_deaths[i]>x$Min_sum_deaths[i]){
    #print(paste("-5-", i, redist))
        x$Impute_deaths[i] <- x$Impute_deaths[i]-1
        redist <- redist - 1
        }
    }
    i <- ((i) %% nrow(x)) + 1
  }
  x <- x %>% arrange(ID) %>% 
    select(-ID)
  df[is.na(df$Deaths),] <- x # add back only rows updated
  df$Impute_deaths
}

foo <- DF_Impute_Deaths_County %>% 
  mutate(adjust=Impute_deaths-Min_sum_deaths) %>% 
  group_by(Condition) %>% 
    mutate(New_impute=Fix_minimum(tibble(Condition,
                                         Deaths, 
                                         Impute_deaths, 
                                         Min_sum_deaths, 
                                         adjust))) %>% 
  ungroup()

###############################   stopped here
#   Add in the previously imputed values of deaths per county per condition
#   summed over the year range

Df_County_yr_imputed <- left_join(Df_County_yr, 
                                  DF_Impute_Deaths_County,
                                  by=c("County", "Condition")) %>% 
  select(-Deaths.y) %>% 
  rename(Deaths=Deaths.x, Imputed_deaths_sum=Impute_deaths)






```



```{r Keep as examples}
###########     Read in By Race and County

Df_race <- read.delim(paste0(path, "TableSheet_crosstab(1).csv"),
                   sep="\t",
                   fileEncoding="UTF-16LE",
                   na.strings="---",
                   skip=1,
                   header=FALSE,
                   colClasses="character",
                   stringsAsFactors = FALSE
                   ) %>% 
  rename("County"=V1, "Condition"=V2, "Race_Black"=V3, "Race_Hispanic"=V4, 
         "Race_Other"=V5, "Race_White"=V6
         ) %>%  
  #   F&^%ers put commas in the numbers
  mutate_at(vars(contains("Race_")), 
         funs(str_remove_all(., ","))) %>% 
  #   Make numbers numbers
  type_convert(col_types = paste0("cc", strrep("i",4)))

###########     Read in By Sex and County

Df_sex <- read.delim(paste0(path, "TableSheet_crosstab(2).csv"),
                   sep="\t",
                   fileEncoding="UTF-16LE",
                   na.strings="---",
                   skip=1,
                   header=FALSE,
                   colClasses="character",
                   stringsAsFactors = FALSE
                   ) %>% 
  rename("County"=V1, "Condition"=V2, "Sex_female"=V3, "Sex_male"=V4 
         ) %>%  
  #   F&^%ers put commas in the numbers
  mutate_at(vars(contains("Sex_")), 
         funs(str_remove_all(., ","))) %>% 
  #   Make numbers numbers
  type_convert(col_types = paste0("cc", strrep("i",2)))

###########     Read in By Total and County

Df_Total <- read.delim(paste0(path, "TableSheet_crosstab(3).csv"),
                   sep="\t",
                   fileEncoding="UTF-16LE",
                   na.strings="---",
                   skip=1,
                   header=FALSE,
                   colClasses="character",
                   stringsAsFactors = FALSE
                   ) %>% 
  rename("County"=V1, "Condition"=V2, "Total"=V3
         ) %>%  
  #   F&^%ers put commas in the numbers
  mutate_at(vars(contains("Total")), 
         funs(str_remove_all(., ","))) %>% 
  #   Make numbers numbers
  type_convert(col_types = paste0("cc", strrep("i",1)))






```














